---
title: examining the data 
output: 
  html_document:
    keep_md: TRUE
---

### read the data file 

If you haven't already, install the *readr* package from RStudio using *Packages -> Install -> readr*. 

Add this text to your Rmd file. Save and Knit. 

```{r include = FALSE}
# match the knitr root directory to the project working directory
library(knitr) 
opts_knit$set(root.dir = '../')
opts_chunk$set(echo = FALSE, results = 'hide')
```

```{r}
# assign the data to a variable name
library(readr)
wide_data <- read_csv('data/009_wide-data.csv')
```

<pre class="r"><code>```{r include = FALSE}
# match the knitr root directory to the project working directory
library(knitr) 
opts_knit$set(root.dir = '../')
<code>```</code>

Input the data file that was provided. 

<code>```</code>{r}
# assign the data to a variable name
library(readr)
wide_data <- read_csv('data/009_wide-data.csv')
<code>```</code>
</code></pre>

What the script does: 

R:

- `# comment` a hash tag in R denotes a comment 
- `library()` loads a package so we can use its functions 
- `read_csv()` reads a CSV file. The path to the data file (`'data/009_wide-data.csv'`) file is relative to the project working directory  
- `<-` is the R assignment operator 

knitr: 

- We match the working directory of the *knitr* package to the working directory of the project folder to allow us to render the Rmd script (which also executes all the R code) or to run the R code alone (without rendering the script) using *Run All* . 



### data structure

Add the following. Save and Knit.  

```{r}
str(wide_data)
```

<pre class="r"><code>Examine the data.

<code>```</code>{r}
str(wide_data)
<code>```</code>
</code></pre>

What the script does: 

- `str()` returns the structure of the R object. The output printed to your Viewer screen shows the object `wide_data` is a data frame having `r nrow(wide_data)` observations (rows) of `r ncol(wide_data)` variables (columns) 

### data frame characteristics

Add the following. Save and Knit. 

```{r}
# data frame dimension
dim(wide_data)

# column names
names(wide_data)

# first few rows
head(wide_data, n = 3L)

# last few rows
tail(wide_data, n = 3L)

# descriptive statistics of each column
summary(wide_data)
```

<pre class="r"><code>Look at the characteristics of the data frame. 

<code>```</code>{r}
# data frame dimension
dim(wide_data)

# column names
names(wide_data)

# first few rows
head(wide_data, n = 3L)

# last few rows
tail(wide_data, n = 3L)

# descriptive statistics of each column
summary(wide_data)
<code>```</code>

We see from this initial exploration that the data consist of `r dim(wide_data)[1]` measurements per cycle with `r length(grep('cycle', names(wide_data)))` cycles. Some data are missing from some cycles (indicated by NA)  which is consistent with the ANSI/ISA standard by which the data were obtained. 
</code></pre>

What the script does:

- `dim()` returns the dimensions of the object 
- `names()` returns the columns names 
- `head()` and `tail()` return the first few and last few rows of the data frame. The number of rows is given by the `n =` argument. 
- `3L` declares the number 3 as an integer (instead of a floating point number) 
- `summary()` returns descriptive statistics for every column 



