---
title: examining the data 
tagline: load cell calibration 
output: 
  html_document:
    keep_md: TRUE
---

```{r include = FALSE}
library(knitr) 
opts_knit$set(root.dir = '../')
```

### instructions

In the pre-workshop homework you installed the *readr* library. If not, the instructions are repeated here: 

- In RStudio, lower right pane, select Packages -> Install 
- In the dialog box, type `readr` 
- Check the box to install dependencies 
- Select Install 

In your Rmd file 

- Add a new section heading `## Examine the data` 
- add code (shown boxed) in code chunks 

Code results (which should also show up in your Viewer pane) are shown boxed with a double hash tag 


### examine the data 

Use the *readr* package *read_csv()* function to bring the calibration data (the data you downloaded earlier) into the R workspace. In a code chunk, add 

```{r}
library(readr)
wide_data <- read_csv('data/009_wide-data.csv')
```


- *library()* loads a package so we can use its functions 
- *read_csv()* reads a CSV file 

> Examine the data: 

- *str()* is an R function that describes the structure of its argument

```{r}
str(wide_data)
```

- *dim()* is an R function that returns the row by column dimensions of the data frame

```{r}
dim(wide_data)
```

- *names()* is an R function that returns the columns headings in a data frame

```{r}
names(wide_data)
```

- *head()* and *tail()* are R functions for examining the first *n* rows and the last *n* rows of a data frame. `3L` is R-syntax declaring that 3 is an integer, not a floating-point number.

```{r}
head(wide_data, n = 3L)
tail(wide_data, n = 3L)
```

- *summary()* provides the descriptive statistics of each column 

```{r}
summary(wide_data)
```

- the data set is small enough in this case that we can print the full set to the output document using *print()*  

```{r}
print(wide_data)
```

> We see from this initial exploration that the data consist of `r dim(wide_data)[1]` measurements per cycle with `r length(grep('cycle', names(wide_data)))` cycles. Some data are missing from some cycles (indicated by NA)  which is consistent with the ANSI/ISA standard by which the data were obtained.
> 
> The data are in *wide format* making them easy to read but difficult for analysis. Reshaping the data from wide format to long format is the next topic.

A reminder, in your scripts, you should: 

- Type (or copy and paste) the R code chunks in full
- Type (or copy and paste) enough of the prose to explain the analysis.
