---
title: "calibration graph"
output:
  html_document:
    keep_md: yes
---

```{r setup, include = FALSE}
library(knitr)
opts_knit$set(root.dir = "../")
opts_chunk$set(echo = TRUE, include = TRUE, eval = TRUE, results = 'hide', message = FALSE, warning = FALSE, fig.keep='none')
```

```{r, echo = FALSE}
# my tutorial functions for text_icon and code_icon
source("scripts/helper_02_icons.R")
```

Packages used in this tutorial  

- readr
- ggplot2

How to use this tutorial 

- `r text_icon` *add text*: type the prose verbatim into the Rmd file 
- `r code_icon` *add code*: insert a code chunk then transcribe the R code 
- `r knit_icon` *Knit* after each addition. 

### open a new Rmd script

Open a new Rmd file, and name it `03_calibr-graph.Rmd`. Save it to the `scripts` directory.

Edit the YAML header:  

    ---
    title: "Load-cell calibration --- graph"
    author: "your name"
    date: "date"
    output: html_document
    ---

Insert knitr setup code

`r code_icon` 

    library(knitr)
    opts_knit$set(root.dir = '../')
    opts_chunk$set(echo = TRUE)

Knowing the packages we'll be using, we can load them right away, near the top of the file.

`r code_icon`

```{r}
# load packages
library(readr)
library(ggplot2)
```

You will most likely encounter three main graphical systems in R: `base graphics`, `lattice`, and `ggplot2`. We'll use `ggplot2`, written by the [nerd famous Hadley Wickham](http://priceonomics.com/hadley-wickham-the-man-who-revolutionized-r/), and currently the most popular R graphics system. 

You should learn to recognize syntax from the other systems too. With so many code samples online, you will eventually want to copy someone's code as a basis for a new graph, and you can't use, say, `lattice` syntax in a `ggplot2` environment. To illustrate the difference in the three systems, scatter plots are drawn using: 

- `plot(x, y, ...)` in `base`
- `xyplot(y ~ x, ...)` in `lattice` 
- `geom_point(aes(x, y, ...))` in `ggplot2`

Robert Kabacoff, the author of the online reference [Quick-R](http://www.statmethods.net/) and the book [R in Action  2/e](https://www.manning.com/books/r-in-action-second-edition) (if you buy one book to help you learn R, I recommend this one), summarizes R graphics systems [here](http://www.slideshare.net/kabacoff/r-for-data-visualization-and-graphics-31577136?next_slideshow=1). Also, an anonymous blogger compares lattice to ggplot2 in a series of posts [here](https://learnr.wordpress.com/2009/08/26/ggplot2-version-of-figures-in-lattice-multivariate-data-visualization-with-r-final-part/). 

Having used `lattice` for years now, I am a novice `ggplot2` user too, but I'm convinced it's the package to learn --- all the cool kids are using it. 

### data

`r text_icon`

    # Data
    
    Read the tidy data set.

`r code_icon`

```{r}
graph_data <- read_csv("data/02_calibr_data-tidy.csv")
str(graph_data)
```

### check yourself

Confer with a neighbor.

1. How many variables in this data frame?
2. Which variables do we use in the calibration graph? 

### building the graph in layers 
    
`r text_icon`

    # Building the graph in layers 
    
    Just show the data. 

`r code_icon`

```{r}
calibr_graph <- ggplot(data = graph_data, aes(x = input_lb, y = output_mV)) +
    geom_point()
``` 

Learning ggplot2

- `data = ...` brings in the data frame
- `aes()` is the **aes**thetic mapping, indicating which of the variables in the data frame are used for `x` and `y` coordinates of the scatter plot 
- `+` is the syntax for "add the next layer"
- `geom_point()` is a **geom**etric object. We use points because we're drawing a scatterplot

`r code_icon`

```{r eval=FALSE}
print(calibr_graph)
```

That's all it takes to obtain a basic scatterplot. If you don't care for the default graphic settings, don't worry, I'll show you how to edit basic settings. 

`r text_icon`

    Draw the regression line.  

`r code_icon`

```{r}
calibr_graph <- ggplot(data = graph_data, aes(x = input_lb, y = output_mV)) +
    geom_smooth(method = 'lm') + 
    geom_point()
```

Learning ggplot2

- `geom_smooth()` computes and draws a fitted curve and confidence interval
- `method` assigns the type of curve fit, `lm` is a linear model
- Layers are drawn consecutively, so by drawing `geom_smooth()` before `geom_point()`, the points overprint the line. 

`r code_icon`

```{r eval=FALSE}
print(calibr_graph)
```

### check yourself

Confer with a neighbor

1. Explain why `geom_line()` and  `geom_point()` have different `aes()` arguments.  

### changing the attributes of graphical elements

`r text_icon`

    # Changing the attributes of graphical elements 
    
    Start with the line and the data markers.

`r code_icon`

```{r}
calibr_graph <- ggplot(data = graph_data, aes(x = input_lb, y = output_mV)) +
    geom_smooth(method = 'lm', se = FALSE, color = 'gray70',  size = 0.5) + 
    geom_point(size = 1.5, stroke = 0.7, shape = 21, color = 'black', fill= 'gray70')
```

Learning ggplot2

- `se = FALSE` turns off the confidence interval band
- The line is edited in size (line width in mm) and color. 
- The point is edited in size (diameter in mm),  [shape](http://docs.ggplot2.org/current/vignettes/ggplot2-specs.html), stroke and color (outer circle), and fill color. 
- R has `r length(colors())` built-in named colors such as `gray70` and `black` I've used here. You can see the full list of color names by typing `colors()` in the Console. View the colors by name  [here](http://www.stat.columbia.edu/~tzheng/files/Rcolor.pdf). 

`r code_icon`

```{r eval=FALSE}
print(calibr_graph)
```

`r text_icon`

    Edit the axis labels. 

`r code_icon`

```{r}
calibr_graph <- calibr_graph + 
	xlab("Applied force (lb)") + 
	ylab("Sensor reading (mV)")
```

Learning ggplot2

- This chunk could be read "Assign `calibr_graph` to itself, add a layer for the x-axis label, and add a layer for the y-axis label."

`r code_icon`

```{r eval=FALSE}
print(calibr_graph)
```

`r text_icon`

    Edit the locations of the axis markings to match the test points. 

`r code_icon`

```{r}
calibr_graph <- calibr_graph + 
	scale_x_continuous(breaks = seq(from = 0.5, to = 4.5, by = 1)) +
	scale_y_continuous(breaks = seq(from = 10, to = 90, by = 20))
```

Learning R

- `breaks` is ggplot2 syntax for tick mark locations
- Sequences have the form `seq (from = ..., to = ..., by = ...)` as shown here. You can shorten the code by omitting the *key* words as long as the *values* are in the order shown. For example, I could have written `seq(10, 90, 20)` and obtained the same result, the vector `(10, 30, 50, 70, 90)`. 

`r code_icon`

```{r eval=FALSE}
print(calibr_graph)
```

`r text_icon`

    Assign a different theme. 

`r code_icon`

```{r}
calibr_graph <- calibr_graph + 
	theme_light()
```

Learning ggplot2

- A theme, in ggplot2, is a collection of settings that controls the appearance of the graph. Example images of the built-in themes are [here](http://docs.ggplot2.org/current/ggtheme.html). 
- Additional themes are available in the [ggthemes package](https://github.com/jrnold/ggthemes), including themes inspired by Stephen Few, Edward Tufte, The Economist magazine, and even the classic 2003 ugly gray charts in Excel ("for ironic purposes only").  

`r code_icon`

```{r eval=FALSE}
print(calibr_graph)
```

`r text_icon`

    Edit the theme.  

`r code_icon`

```{r}
calibr_graph <- calibr_graph +
		theme(panel.grid.minor = element_blank(),  
					axis.ticks.length = unit(2, "mm"))
```

Learning ggplot2

- `element_blank()` turns off the minor grid lines
- `axis.ticks.length` edits the length of the axis tick marks

The ability to edit or create a theme is the essential tool for customizing graphs in `ggplot2`. I've show two small customizations; the full list is given [here](http://rstudio-pubs-static.s3.amazonaws.com/3364_d1a578f521174152b46b19d0c83cbe7e.html).  

`r code_icon`

```{r eval=FALSE}
print(calibr_graph)
```

### putting it all together 

`r text_icon`

    # Putting it all together
    
    We've written out the layers in separate code chunks for clarity, but they can be assembled into one code chunk in the same order as before. A design feature of  `ggplot2` syntax is that a graph is constructed one layer at a time---and the structure of the layers is revealed by the structure of the code.  
    
`r code_icon`

```{r}
calibr_graph <- ggplot(graph_data, aes(input_lb, output_mV)) +
    geom_smooth(method = 'lm', se = FALSE, color = 'gray70',  size = 0.5) + 
    geom_point(size = 1.5, stroke = 0.7, shape = 21, color = 'black', fill= 'gray70') +
    xlab("Applied force (lb)") + 
    ylab("Sensor reading (mV)") +
    scale_x_continuous(breaks = seq(0.5, 4.5, 1)) +
    scale_y_continuous(breaks = seq(10, 90, 20)) +
    theme_light() +
    theme(panel.grid.minor = element_blank(), axis.ticks.length = unit(2, "mm"))
```

Learn ggplot2. Earlier, we shortened the `seq()` function by writing the arguments in a particular order. We can do the same with other functions: 

- From `ggplot()`, you can omit `data = `, leaving the data frame name
- From `aes()`, you can omit `x = ` and `y = `, leaving the x-variable and y-variable names in that order

`r code_icon`

```{r eval=FALSE}
# print to screen
print(calibr_graph)
```

And, print to file. 

`r code_icon`

```{r}
# print to file
ggsave("results/03_calibr_graph.png", plot = calibr_graph, 
			 width = 6, height = 4, units = "in", dpi = 300)
```

Learning ggplot2

- Save to the `results` directory because the graph will be used in a report
- `dpi = 300` for using in a print document
- `dpi` $\times$ `width` and `height` in inches yields screen dimension in pixels (72 ppi for web resolution is a  [myth](http://www.photoshopessentials.com/essentials/the-72-ppi-web-resolution-myth/)).


### check yourself

Navigate to your scripts and results directories. They should look like this:

    scripts\
      |-- 01_calibr_data-reshaping.Rmd 
      |-- 02_calibr_data-tidying.Rmd
      `-- 03_calibr_graph.Rmd
      
    results\
      |-- 02_calibr_data-tidying.csv
      `-- 03_calibr_graph.png
      
      
      

### advanced graphics

There's much more we can do to bring the graph up to publication standards. 

I've prepared a (completely optional) tutorial that develops some [graph extras](116_graph-extras.html)  for those of you who want to know MORE and you want it NOW.  Topics include: 

- Jitter the data markers to reduce printing overlap and set a random seed so jittering is repeatable 
- Use the results table to dynamically assign units to the axis labels and set  tick mark locations
- Use the results table to dynamically edit the calibration equation and print it on the graph
- Format the calibration equation for significant figures
- Change font sizes



      
---
Back [data tidying](110_data-tidying.html)<br>
Next [linear regression](112_linear-regression.html)

